ExEl: Building an Elaborator Using Extensible Constraints

### 2024-03-04 paper/main.tex
intro and abstract attempt #1

Dependently typed languages such as Agda, Coq, Lean, and Idris are used for statically enforcing properties of programs and for enabling type-driven development.
While there has been a lot of work to find the right theoretical foundations for the core type theory underlying these languages, their implementations have been studied to a smaller extent.
In particular, theoretical works rarely consider the plethora of features that exist in big languages like Agda.
As a consequence the developers have to make a lot of adaptations in the implementations, which leads to codebases that are hard to maintain.
In this work we focus on a part of the type-checker called the elaborator that transforms user-friendly surface syntax into a small well-behaved core theory.

@@

This allows for a more compact base elaborator implementation while enabling extensions to the type system.

@@

@@

  pdfcreator={LaTeX via pandoc}}

>>>>

  There has been a lot of work to find the right theoretical foundations for the core type theory underlying proof assistants and dependently-typed programming languages like Coq, Agda, Lean or Idris.
  However, their implementations have been studied to a smaller extent.
  Together with the additions of features that aren't reflected in theoretical studies, this means that developers have to make a lot of adaptations in the implementations, making the codebases that are hard to maintain.
  In this work we consider some common problems in the implementations of the type-checkers and propose a new architecture that allows for a more modular development and containment of features to one module each.
  We focus focus on a part of the type-checker called the elaborator that transforms user-friendly surface syntax with a lot of features into a small well-behaved core theory.

@@

  We showcase the design on a proof-of-concept implementation.

@@

  pdfauthor={Bohdan Liesnikov and Jesper Cockx},
  pdfkeywords={dependent types, elaborators, elaboration, open datatype, extensible datatype, type-checking},
  pdfcreator={LaTeX via pandoc},
  pdfproducer={pdftex},

@@

  }
